%==============================================================================
\documentclass[handout,english]{beamer}
%\documentclass{beamer}
%==============================================================================
\usepackage{epigraph}
\usepackage{longtable}
\usepackage{pifont}
\usepackage{media9}  
\usepackage{Preamble_files/beamer_preamble}
\usepackage{appendixnumberbeamer}
\usepackage{comment}
\usepackage{RKsmileys}
\setbeamertemplate{itemize item}{
  \scriptsize\raise1.25pt\hbox{
    \donotcoloroutermaths$\blacktriangleright$
  }
}
\setbeamertemplate{itemize subitem}{
  \tiny\raise1.5pt\hbox{
    \donotcoloroutermaths$\blacktriangleright$
  }
}

\def\checkmark{
  \tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;
}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\linespread{1.5}
%\input{../../bib/_addbibresource_Alphabetical_Aligned.tex}

\usetheme{}
\title{Sustainability, Modelling and Regional Transition (in Queensland)}
\author[Patrick O'Callaghan]{Patrick O'Callaghan, John Mangan, Tina Rampino}
\institute{AIBE, The University of Queensland\\
\today}
\date{}
\AtBeginSection[]
{
  \begin{frame}{Presenters}\thispagestyle{empty}
    \addtocounter{framenumber}{-1}
\tableofcontents[currentsection]
\vfill\end{frame}
\normalsize
}
\AtBeginSubsection[]
{
%\begin{frame}{Table of Contents}
%\vskip8pt
%\tableofcontents[currentsection, currentsubsection]
%\vfill\end{frame}
%\normalsize
}


\begin{document}
\maketitle
\section{John Mangan}
\begin{frame}
%==============================================================================
  \frametitle{Why are we here?}
%==============================================================================
  Brief discussion of the history of the project and why we are here.
  \begin{itemize}
    \item why the project is important to AIBE\\
    \item background of regional/sectoral economics at UQ\\
    \item Input-Output modelling\\
    \item the state of CGE modelling in Australia\\
    \begin{itemize}
      \item CoPS, U Victoria (no uncertainty at all or proper dynamics)\\
    \end{itemize}
  \end{itemize}
\end{frame}
\section{Patrick O'Callaghan}
\begin{frame}
%==============================================================================
  \frametitle{Purpose}
%==============================================================================
  Yesterday in conversations with Flavio and John:
  \begin{itemize}
    \item Flavio: CGE is good for long-term problems
    \item John: real test of success is if we can provide new insights
  \end{itemize}
  Key points:
  \begin{itemize}
    \item Maiwar expands the scope of CGE to mid-term problems
    \item Tools we now have are being applied to design electricity grids
\end{itemize}
\end{frame}
\begin{frame}
%==============================================================================
  \frametitle{How we arrived at Maiwar: methodology}
%==============================================================================
  Recent literature on macroeconomic production networks: Baqaee--Farhi, \dots
  \begin{itemize}
    \item embeds network modelling in macro models (IO renaissance)
    \item main finding: nonlinear (beyond Cobb-Douglas) effects matter
  \end{itemize}
  Atalay has some good econometric estimates of elasticities
  \begin{itemize}
    \item elasticity of substitution for flows between sectors is about 0.1
    \item previously most assumed Cobb-Douglas elasticity of 1.0
  \end{itemize}
  Conclusion: nonlinear effects matter
  \begin{itemize}
    \item {\color{patrickcolor3}
      does not bode well for multi-sectoral steady-state approx.}
    \item this includes Baqaee--Farhi, Atalay, CoPS, McKibben
  \end{itemize}
\end{frame}
\begin{frame}
%==============================================================================
  \frametitle{How we arrived at Maiwar: methodology}
%==============================================================================
  How about ``global'' approximation e.g. Value Function Iteration?

  Building on Scheidegger's machine-learning approaches, we found
  \begin{itemize}
    \item for 1-sector model, grid approach is (at least) 1 order of magnitude 
      more accurate than steady/ergodic-state approach
  \end{itemize}
  But, after a lot of hard work:
  \begin{itemize}
    \item VFI is also unstable (poorly suited to multi-sectoral flows)
    \item trouble with grid approach: $10 ^ {12}$ points for 12 dimensions
    \item approximation is poor outside of grid
  \end{itemize}
\end{frame}
\begin{frame}\thispagestyle{empty}
%==============================================================================
  \frametitle{MAIWAR model literature}
%==============================================================================
  \begin{itemize}
    \item Solution method: Cai--Judd (2021, 2017)
    \item Model is a combination of Cai--Judd and Atalay
    \item Calibration/estimation methods and estimates from Atalay
    \item But Atalay's solution is not dynamic/stochastic
      \dots
    \item We next explain Cai--Judd in some detail.
  \end{itemize}
\end{frame}
\begin{frame}
  {\color{blue} Quote from Cai and Judd (Feb, 2021)}
  \begin{quote}
  {\color{blue} Macroeconomists [and CoPS] are often interested in obtaining solutions around the non-stochastic steady state.}
  {\color{patrickcolor3} However in reality, the initial state could be far away from the steady state, and a policymaker may be more interested in the solutions for the initial periods in the forward-looking model than the far future states that could be around the steady state.}
  {\color{patrickcolor1} For example, in environmental and climate change economics \dots}
\end{quote}
\end{frame}
\begin{frame}
%==============================================================================
  \frametitle{A bit more detail on the current state of CGE modelling}
%==============================================================================
  \begin{itemize}
      \item McKibbin 1990s GCube (steady-state, no uncertainty either)
      \item Malakellis 1990s CoPS (Hamiltonian no uncertainty)
      \item Ha--Kompas 2017 Hamiltonian, no uncertainty, impressive.
      \item Wende 2019, Treasury Intersectoral Model (TIM) steady-state,
        no uncertainty
  \end{itemize}
\end{frame}
\begin{frame}
%==============================================================================
  \frametitle{How we arrived at Maiwar \dots the research question}
%==============================================================================
  Herakles Polemarchakis: building a model without a question is 
  \emph{putting the cart before the horse}.

  In 2022, settled on 2050 net-zero carbon emission targets.
    \begin{itemize}
      \item Most economists think Australia will be better off, but
      \item Adams 2021, CoPS: Qld -6\% GSP and -100k jobs \emph{rel. to base}.
      \item What about Qld targets over and above those of Australia?
    \end{itemize}
\end{frame}

\begin{frame}
%==============================================================================
  \frametitle{MAIWAR \small(Modelling Australian Industry: Weak-Solutions, AMPL,
  Regions)}
%==============================================================================
  \begin{enumerate}\footnotesize
    \item Flexible yet Fast: without steady-state approx
    \item Look-forward property: flow of state-action dependent rewards
  \[
    V_{t_0}(\omega_{t_0}) = r_{t_0}(\omega_{t_0}, a_{t_0}) + \cdots
    + r_{t_9}(\omega_{t_9}, a_{t_9}) + V_{t_{10}}(\omega_{t_{10}})
  \]
    \item Uncertainty: easy way to improve on CoPS
    \item Investment/saving behaviour: Euler equations: CGE Dixon--Rimmer 2020
    \item Data: BLADE, calibration, econometrics
    \item Robust/Reliable: works with a variety of set-ups
    \item Accurate/Accessible: as open source as possible
    \item Modern yet Trustworthy: best-in-class knowledge, 2+ solvers
    \item Scalable: at least to 8 regions and 20 sectors
  \end{enumerate}\normalsize
\end{frame}
\begin{frame}
%==============================================================================
  \frametitle{Maiwar: the current state of affairs}\thispagestyle{empty}
%==============================================================================
  Currently developing strategies for easy Scaling (pre Calibration).
    \begin{itemize}
      \item Conopt: for a well-scaled model, all values (variable,
        dual variable and derivative) in $[10^{-2} , 10^{+2}]$.
      \item E.g. Beyond 10 sectors log Cobb-Douglas Utility:
        \emph{corner sol}.
        \begin{itemize}
          \item To help identify scaling parameters: NIMROD, RCC working!
          \item Last Friday: 120 runs in 1 hour on 12 CPUs.
        \end{itemize}
    \end{itemize}

  In February, we obtained access to BLADE and established that can use for
  for government contracts (\emph{not for commercial gain}).
  \begin{itemize}
    \item Illion Credit-reporting agency: possible BLADE integration \dots
  \end{itemize}
\end{frame}
\begin{frame}
%==============================================================================
  \frametitle{Quick interlude on formal model}
%==============================================================================
\uncover <+-|alert@+>{\color<.>{blue}
  \begin{align*}
    V_{0}(\omega_{0}) = \max_{a_{\cdot}}& \quad \mathbb E
    \left\{\sum_{t = 0}^{9} \beta^{t} r_{t}(\omega_{t}, a_{t}) + \beta^{10} V_{10}(\omega_{10})\right\}
    \\
    \textup{s.t.}& \quad \omega_{t+1} = g_{t}(\omega_{t}, a_{t}, \varepsilon_{t}), \quad t = 0, \dots, 9
    \\
    & \quad f_{t}(\omega_{t}, a_{t}) \geq 0, \quad t = 0, \dots, 9.
  \end{align*}
}
\uncover<+-|alert@+>{\color<.>{blue}
With current state $\omega_{t}$, action  $a_{t}$, expectation $\mathbb E$, reward $r_{t}$, discount factor $0 < \beta < 1$, terminal value function $V_{10}$, transition law $g_{t}$, error $\varepsilon_{t}$ and feasibility constraints $f_{t}$ on actions.}
\\
\begin{itemize}
  \item Grid-free: all about paths \dots a simple example.
\end{itemize}

\vfill\end{frame}
\begin{frame}
%==============================================================================
  \frametitle{Cai--Judd's SCEQ: \small Simple (yet Powerful) Certainty
  Equivalent Method}
%==============================================================================
  E.g. Irreversible risk:  one-off, permanent $5\%$ productivity shock.

  Loosely resembles tipping points: each year, chance of ice-shelf \dots

  In this simple example there are 28 paths to 2050:
    \begin{itemize}\small
    \item path where shock never happens
    \item path where shock happens in 2023;
    \item path where shock happens in 2024;
    \item \dots
    \end{itemize}\normalsize
  Each path has 28 periods (PathTimes).

  Agents make 10-year plans (LookForward) at each time and path.
  
  \small
  Goal: Balance consumption today vs uncertain consumption tomorrow.
  On path $p$: once plan $t_0$ is made, time reveals state $\omega_{t_{1}}$
 $\rightarrow$ new plan.

\end{frame}
\begin{frame}
%==============================================================================
  \frametitle{What do we (the modeller) do with all this info?}
%==============================================================================
  \begin{itemize}
    
    \item For each path $p$ and step $s$ along the path, take the plan that
      starts at time $s$ and store the values for time $s$.
    \begin{itemize}
      \item For consumption store in a matrix $C_{s, p}$. The values of this
        matrix are the solutions for each $r$ in Regions and $i$ in Sectors.

      \item similarly for investment $\textup{INV}_{s, p}$, labour $L_{s, p}$,
        kapital $K_{s, p}$, \etc.
    \end{itemize}
    \item \emph{We have generated an empirical distribution of solutions.}

    \item Can now derive the sample means (paths), variances, \etc.

    \item Check that the Euler equation holds for the sample mean.

  \end{itemize}
\end{frame}
\begin{frame}
%==============================================================================
  \frametitle{Example of results}
%==============================================================================
  
  See spreadsheet and live-run output.
\end{frame}
\begin{frame}\thispagestyle{empty}
%==============================================================================
  \frametitle{Main contribution to Cai--Judd: multi-sectoral flows}
%==============================================================================
  For each Region $r$, Sector $j$ and LookForward time $t$:
  \[
    k_{r, j, t + 1} = (1 - \delta) k_{r, j, t} + s_{r, j, t}
  \]
  where $s_{r, j, t}$ is a CES function of intermediate Long--Plosser flows.
  \[
    s_{r, j, t} = \left(\sum_{i} \sigma_{i j} S_{r, i j, t} ^ {\rho} \right)
      ^ {\frac{1}{\rho}}
  \]
  From Atalay's model: $\rho = \frac{0.1 - 1}{0.1} = -9$.

  But: $20$ Sectors implies $400$ flows: $8 \times 400 \times 10 = 32,000$ in 
  total.

  Jacobi Equations, for $\mathbf{S}_{r, ij, t} = S_{r, ij, t} / \sigma_{r, ij}$:
    \[
      \mathbf{S}_{r, ij, t} = \mathbf{S}_{r, ik, t} \times
      \mathbf{S}_{r, kj, t}^{-1} \times \mathbf{S}_{r, kj, t} .
    \]
  Use main diagonal and one column of $S$: $39$ flows with total $3,120$.
\end{frame}
\begin{frame}
%==============================================================================
  \frametitle{Other innovations}
%==============================================================================
So far, the main contributions are:

To CGE: a tractable way of modelling uncertainty.

To Atalay: we take his model beyond the steady-state.

To Cai--Judd: a framework for sectoral/production networks.

\end{frame}
\begin{frame}
%==============================================================================
  \frametitle{Cai--Judd: under the bonnet}
%==============================================================================
  For each plan, the terminal value function $V_{10}$ is arbitrary.

  One solution is to make longer plans: discount factor 
    $\beta^t \rightarrow 0$.
    \begin{itemize}
      \item Behaviourally unrealistic?
      \item Costly computationally \dots
    \end{itemize}
  So best to treat it as another form of parametric uncertainty.
    \begin{itemize}
      \item Scaling this term: important for avoiding corner solutions.
    \end{itemize}
  Checking the Euler Equations in multi-sectoral models is tricky compared to
  Cai--Judd baseline models.
    \begin{itemize}
      \item Maximum of deviations from the Jacobi Identity helps here.
        \begin{itemize}
          \item my first guide to a good solution.
        \end{itemize}
    \end{itemize}
\end{frame}
\begin{frame}
%==============================================================================
  \frametitle{Data}
%==============================================================================
  Late December 2020: lunch with Alicia, discussion of regional
  identifiers that had only just been added to BLADE.
  \begin{itemize}
    \item This represents a unique opportunity for regional modelling
    \item Moreover, Type of Activity Units
      \begin{itemize}
        \item ABS uses employment data to split larger firms
      \end{itemize}
    \item This means upto SA1-level regions and 4-digit ANZSIC sectors!
  \end{itemize}

\end{frame}

\section{Tina Rampino}
\begin{frame}\small
%==============================================================================
  \frametitle{BLADE: \small 	Business Longitudinal Analysis Data Environment} 
%==============================================================================
  \begin{itemize}
    \item Economic data tool integrating tax, trade and intellectual property data 
  with information from ABS surveys
    
   \item Contains data on all active businesses from 2001-02 to 2018-19.
	
   \item We are currently using BLADE Core to construct parameters to scale down
  the Input-Output tables for Australia.

   \item Approx. 2.4 million operating businesses in 2020
  
   \item Patrick and I currently have access and are building 
  variables that will allow us to capture number of TAUs in a given region.

   \item Alicia and Martie-Louise are discussants.
\end{itemize}
\end{frame}
\begin{frame}\small
%==============================================================================
  \frametitle{BLADE benefits and usage}
%==============================================================================
\begin{itemize}
	  \item Publicly-available data not granular enough for local policy makers. 
	  \item BLADE is especially relevant in the case of Qld (state with the highest regional population)
	  \item Projects must benefit the public:
	  \item SMaRT in Qld modelling provides insights on policy interventions
	  \item “Pulse of the region” industry engagement meetings: DAF interested in assessing the impact of COVID-19 using CGE framework (they think SA4 level, we can model SA1, which is more relevant for rural Queensland)
  \end{itemize}
\end{frame}
\section{Other team members and thanks}
\begin{frame}
%==============================================================================
  \frametitle{Thanks for listening!}
%==============================================================================
\uncover<+-|alert@+>{\color<.>{blue}
And many thanks to the absent members of the team:
}

\begin{itemize}
\item<+-|alert@+>{\color<.>{blue}
        Josh Aberdeen (UQ Student Joint Bachelors)}
\item<+-|alert@+>{\color<.>{blue}
        Patrick Duenow (UQ Student PhD Health Economics)
        }
\item<+-|alert@+>{\color<.>{blue}
        Cameron Gordon (U Adelaide Student PhD Machine Learning)
        }
\item<+-|alert@+>{\color<.>{blue}
    Alexander Yelkhovsky (Research Fellow Computing)
  }
\end{itemize}
Also Alicia Rambaldi for her advice on BLADE

Jorge Miranda-Pinto for his advice and expertise on the literature

Brent Ritchie for making the project happen.

\vfill\end{frame}
%==============================================================================
\section{Appendix/extra slides}
%==============================================================================
\begin{frame}
%==============================================================================
  \frametitle{Why CGE modelling?}
%==============================================================================

  \begin{itemize}
    \item<+-|alert@+>{\color<.>{blue}
      "Old-fashioned", "black-box", "intractible", \dots
    }
    \item<+-|alert@+>{\color<.>{blue}
      Yet industry demands "CGE" modelling and uses it as a basis for key 
      decisions.
    }
    \item<+-|alert@+>{\color<.>{blue}
      A guide to quantifying the broader repurcussions of sector-specific
      shocks.
    }
    \item<+-|alert@+>{\color<.>{blue}
      A guide to analysing the sector-specific implications of macroeconomic shocks
    }
  \end{itemize}
\vfill
\end{frame}
\begin{frame}
%==============================================================================
\frametitle{Why the focus on uncertainty?}
%==============================================================================
John Williams (Federal Reserve, San Fransisco):

\begin{quote}
There is pretty strong evidence that the rise in uncertainty is a significant factor holding back the pace of recovery now. [...] research shows that heightened uncertainty slows economic growth, raises unemployment, and reduces inflationary pressures. [...] There is no question that slow growth, high unemployment, and significant uncertainty are challenges for monetary policy.
\end{quote}
From Caggiano, Castelnuovo, Groshenny 2015.
\end{frame}

%\begin{frame}
%\frametitle{{\color{patrickcolor3}
%  Example: 2050 net-zero emissions targets
%}
%}
%\uncover<+-|alert@+>{\color<.>{blue}
%  In 2017, Qld set a { \color{magenta} \href{https://www.vision6.com.au/em/message/email/view.php?id=1398053&a=79404&k=VZGdbxVB3M5LdOkRxoA2ohzW3lZeME6xaZFD4f5Ic-o}{``state target for net-zero emissions by 2050''}
%  }.
%}
%\begin{itemize}
%\item<+-|alert@+>{\color<.>{blue}
%        Why not simply enforce the Australia-wide net-zero target?
%        }
%\end{itemize}
%
%\uncover<+-|alert@+>{\color<.>{blue}
%        Likely answer: green credentials beyond the national policy \dots
%      }
%\begin{itemize}
%\item<+-|alert@+>{\color<.>{blue}
%        Norwegian Sovereign Fund divestment of Qld public debt?
%        }
%\item<+-|alert@+>{\color<.>{blue}
%        Exports to EU, US, \dots?
%        }
%  \item<+-|alert@+>{\color<.>{blue}
%        Local public opinion?
%        }
%  \item<+-|alert@+>{\color<.>{blue}
%        Encourage innovation \eg\ hydrogen?
%        }
%\end{itemize}
%\uncover<+-|alert@+>{\color<.>{patrickcolor3}
%  What is the  cost of this policy? (A normative question.)
%}
%\begin{itemize}
%
%  \item<+-|alert@+>{\color<.>{blue} {\color{magenta}\href{https://theconversation.com/economists-back-carbon-price-say-benefits-of-net-zero-outweigh-costs-169939}{link}} (to Conversation article)
%        }
%
%%  \item<+-|alert@+>{\color<.>{blue}
%%        We want a model that can (help) answer such questions.
%%        }
%\end{itemize}
%\vfill\end{frame}
%\begin{frame}
%\frametitle{What zoo of models can answer such questions?}
%\uncover<+-|alert@+>{\color<.>{blue}
%}
%\begin{itemize}
%\item<+-|alert@+>{\color<.>{blue}
%  The preferred policy is an economy-wide carbon price.
%        }
%\end{itemize}
%\uncover<+-|alert@+>{\color<.>{patrickcolor3}
%        Answer: economy-wide models (built on normative principles)
%      }
%      \begin{itemize}
%\item<+-|alert@+>{\color<.>{blue}
%        Leontief Input-Output models (1950s+)
%        }
%\item<+-|alert@+>{\color<.>{blue}
%        Centre of Policy Studies (CoPS) CGE model (1970s+)
%              }
%      \end{itemize}
%\uncover<+-|alert@+>{\color<.>{patrickcolor3}
%        Inter-sectoral models with inter-temporal choice and uncertainty:
%      }
%      \begin{itemize}
%        \item<+-|alert@+>{\color<.>{blue}
%          Original model: Long and Plosser (1983)
%        }
%        \item<+-|alert@+>{\color<.>{blue}
%        More recently: Atalay (2017); Baqaee--Farhi \dots
%        }
%        \item<+-|alert@+>{\color{red}
%        Treasury Intersectoral Model (2019)
%        }
%      \end{itemize}
%      \vfill\end{frame}
%\section{Steady-state models}
%\subsection{CoPS}
%
\begin{frame}
{\color{patrickcolor3} Why not use a CoPS CGE model?}
\begin{itemize}
\item<+-|alert@+>{\color<.>{blue}
Cost of software and of data for the model.
}
\vskip5pt
\item<+-|alert@+>{\color<.>{blue}
  CoPS already have a recent paper on 2050 targets
        }
\end{itemize}
\vskip5pt
\uncover<+-|alert@+>{\color<.>{blue}
Moreover:
\begin{itemize}
\item<+-|alert@+>{\color<.>{blue}
        no proper savings/investment: intertemporal behaviour
        }
        \begin{itemize}
\item<+-|alert@+>{\color<.>{blue}
        leads to strange ``macro-closure conditions''
        }
        \end{itemize}
\item<+-|alert@+>{\color<.>{blue}
        no model of risk/uncertainty and associated behaviour
        }
\end{itemize}
}

\uncover<+-|alert@+>{\color{patrickcolor1}
        CoPS assume current economy is in \emph{Deterministic Steady-State}.
      }
      \vskip5pt
\vfill\end{frame}
\begin{frame}
  {Treasury Intersectoral Model (TIM, 2017)
  }

\uncover<+-|alert@+>{\color<.>{blue}
Part of a new generation of Australian models
}
\vskip5pt
\uncover<+-|alert@+>{\color<.>{blue}
J. Miranda-Pinto of UQ had a hand in TIM (and in our choices)
}
\vskip5pt
\uncover<+-|alert@+>{\color<.>{blue}
  TIM has a sister called EMMA (Macro-econometric forecasting)
  }
\begin{itemize}
\item<+-|alert@+>{\color<.>{blue}
        TIM has proper savings
        }
\item<+-|alert@+>{\color<.>{blue}
        114-sector model of Australia
        }
\item<+-|alert@+>{\color<.>{blue}
        but no risk
         }
\item<+-|alert@+>{\color<.>{blue}
        Deterministic steady state \& {\color{red} we can't access}
      }
\end{itemize}
  \end{frame}
\begin{frame}
{Adapting the Atalay model}
  \uncover<+-|alert@+>{\color<.>{blue}
      Atalay assumes economy in (non-determistic) steady state
      }
\begin{itemize}
  \item<+-|alert@+>{\color<.>{blue}
        pretty complete and quite good empirical foundations
        }
  \item<+-|alert@+>{\color<.>{blue}
        {\color{patrickcolor3} we have full access via Matlab \& Stata}
        }
\end{itemize}
\uncover<+-|alert@+>{\color<.>{patrickcolor1}
      Our adaptation of Atalay is the first model in our suite.
    }
\begin{itemize}
  \item<+-|alert@+>{\color<.>{blue}
        less than 1 second to solve a 20-sector model
        }

  \item<+-|alert@+>{\color<.>{blue}
        regionalise using LGA-level income data via Table Builder
        }
  \item<+-|alert@+>{\color<.>{blue}
        capital flows matrix by adapting a US flows table from 1997 \RKconfused\
        }
  \item<+-|alert@+>{\color<.>{blue}
        Social Accounting Matrix using Current and Capital Accounts}
\end{itemize}
  \end{frame}

%
%\subsection{Estimating and approximating Value functions}
%\begin{frame}
%  {Estimating/approximating $V_0$ and $V_{28}$}
%  \begin{itemize}
%    \item<+-|alert@+>{\color<.>{blue}
%          Estimate parameters:
%          }
%          \begin{itemize}
%              \item<+-|alert@+>{\color<.>{blue}
%                  Entity data: Jobs in Australia, Consumption Surveys, BLADE
%                  }
%              \item<+-|alert@+>{\color<.>{blue}
%                  Link data: Input-Output Tables, ABS (Elazar) supply-chain, illion + BLADE
%                  }
%              \item<+-|alert@+>{\color<.>{blue}
%                  OECD, \etc : productivity and emissions data
%                  }
%          \end{itemize}
%  \end{itemize}
%\uncover<+-|alert@+>{\color<.>{blue}
%  Three approaches to {\color{patrickcolor1}``globally''  \footnotesize(not Atalay)} approximating $V_{\cdot}$
%}
%\begin{itemize}
%        \item<+-|alert@+>{\color<.>{blue}
%        Scheidegger--Bilionis: Gaussian Process model
%        \begin{itemize}
%          \item<+-|alert@+>{\color<.>{blue}
%               working Python model on GitHub
%                }
%          \item<+-|alert@+>{\color<.>{blue}
%               robust (uses Ipopt)
%                }
%          \item<+-|alert@+>{\color<.>{blue}
%                slow, but Nimrod of UQ HPC will solve this
%                }
%          \item<+-|alert@+>{\color<.>{blue}
%                scratch/sandbox model in Julia
%                }
%          \end{itemize}
%        }
%        \end{itemize}
%\end{frame}
%
%\begin{frame}
%  {Two more novel approaches}
%\uncover<+-|alert@+>{\color<.>{blue}
%  Azinovic--Gaegauf--Scheidegger: Deep Equilibrium Networks
%}
%  \begin{itemize}
%    \item<+-|alert@+>{\color<.>{blue}
%          {\color{patrickcolor3} full access to Python code}, we have basic model
%        }
%    \item<+-|alert@+>{\color<.>{blue}
%          Flexible: approximate around ergodic set \& easily use GPUs
%          }
%    \item<+-|alert@+>{\color<.>{blue}
%          Robust: involves approximation by composing linear functions
%          }
%        \end{itemize}
%\uncover<+-|alert@+>{\color<.>{blue}
%  Cai and Judd: Simulated Certainty Equivalent Approximation
%  }
%  \begin{itemize}
%    \item<+-|alert@+>{\color<.>{blue}
%          {\color{patrickcolor3} full access to GAMS code}, translating to Julia, uses Ipopt
%          }
%  \item<+-|alert@+>{\color<.>{blue}
%          {\color{patrickcolor1} Made for 2050-style questions}
%          }
%    \item<+-|alert@+>{\color<.>{blue}
%          Fast: grid-free without relying on SS or ergodic set
%          }
%          \begin{itemize}
%    \item<+-|alert@+>{\color<.>{blue}
%          simulates paths of length 28, Nimrod useful again (for larger models)
%                  }
%          \end{itemize}
%        \end{itemize}
%\end{frame}
%\section{Summary}
%\subsection{Table}
%\begin{frame}
%\frametitle{}
%
%\uncover<+-|alert@+>{\color<.>{blue}
%
%
%\begin{longtable}{| c | c | c | c | c |}
%\hline
%& Atalay & GP & DEQN & SCEQ
%\\\hline
%SS & \cmark & \xmark &  \xmark & \xmark
%\\\hline
%  ``Global'' & \xmark & \cmark & \cmark & \cmark
%\\\hline
%Gridfree & \cmark & \xmark &  \cmark$^{*}$ & \cmark
%\\\hline
%Robust &  \cmark & \cmark &  \cmark & \cmark
%\\\hline
%Fast$^{\dagger}$ & \cmark & \xmark$^{\ddagger}$ & \xmark$^{\ddagger}$ & \cmark
%\\\hline
%Complete &  \cmark$^{\mathsection}$ & \cmark$^{\mathsection}$ & \cmark$^ {\mathsection} $ & \cmark$^{\mathsection}$
%\\\hline
%
%\end{longtable}
%}
%\begin{itemize}
%  \item[ $ * $]<+-|alert@+>{\color<.>{blue}
%        optional
%        }
%  \item[$\dag$]<+-|alert@+>{\color<.>{blue}
%                my math will simplify:  $20 \times 20 = 400 \mapsto 39$ dimensions
%                }
%  \item[$\ddagger$]<+-|alert@+>{\color<.>{blue}
%                Nimrod and GPU respectively
%                }
%\item[ $ \mathsection $]<+-|alert@+>{\color<.>{blue}
%        various stages of development
%        }
%
%\end{itemize}
%\vfill\end{frame}



%\printbibliography
\addtocounter{framenumber}{-5}
\end{document}
